Combinatorial Identities from Lagrange's Interpolation Polynomial

### 2021-01-20 spectral radius of matrix new method/thesis/article012121.tex
0121 final check and sending

In \cite{chang}, Cheng and Weng give many bounds of the spectral radius of a nonnegative square matrix. And based on their theory and Perron-Frobenius theorem, we give another approach to obtain an upper bound of the spectral radius and apply it on the adjacency matrix of a directed graph.

@@

\begin{proof}
    Multiplying the nonnegative vector $u$ in assumption~\ref{pre_thm_em2}

@@

%edges
\draw[edge] (1) to (2);
\draw[edge] (1) to[bend left] (3);
\draw[edge] (1) to[bend left] (4);
\draw[edge] (2) to (3);
\draw[edge] (2) to[bend left] (4);
\draw[edge] (3) to[bend left] (1);
\draw[edge] (3) to[bend left] (4); 
\draw[edge] (4)[bend left] to (1); 
\draw[edge] (4)[bend left] to (2); 
\draw[edge] (4)[bend left] to (3); 
\end{tikzpicture}
\end{center}

@@

%edges
\draw[edge] (1) to (2);
\draw[edge] (1) to[bend left] (3);
\draw[edge] (1) to[bend left] (4);
\draw[edge] (2) to (3);
\draw[edge] (2) to (4);
\draw[edge] (3) to[bend left] (1);
\draw[edge] (3) to (2);
\draw[edge] (3) to[bend left] (4);
\draw[edge] (4)[bend left] to (1);
\draw[edge] (4)[bend left] to (3);
\end{tikzpicture}
\end{center}

Note that both $G$ and $G'$ have the same number of edges, 
neither $A$ majories $A'$ nor $A'$ majories $A$. Our main result still can do comparison between 
$\rho(A)$ and $\rho(A')$. 
We first observe that $A$ is irreducible and $A'$ is $k$-rooted for $k=3$. Moreover 
$A'Q=\begin{pmatrix}
    0 & 0 & 1 & 2\\
    1 & 0 & 1 & 1\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 1
    \end{pmatrix}\geq \begin{pmatrix}
    0 & 0 & 1 & 2\\
    1 & 0 & 0 & 1\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 1
    \end{pmatrix}=AQ$. 
Hence Hence $\rho(A)\leq\rho(A')$ by Theorem~\ref{thm:conclusion} with $C=A$ and $C'=A'$. 
Both of the values of $\rho(C)$ and $\rho(C')$ are close to $2.511547$ by  calculating on computer. 
\end{exam}

@@

\begin{exam}
    Consider the following two $4\times 4$ matrices
    $$C=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 0 & 1\\
    1 & 1 & 0 & 0\\
    1 & 1 & 1 & 0
    \end{pmatrix},\quad C'=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 1 &  0\\
    1 & 1 & 0 & 0\\
    1 & 1 & 1 & 0
    \end{pmatrix}.$$
    With $n=4$ and $k=3$, we have  
    $$CQ=\begin{pmatrix}
    0 & 0 & 1 & 2\\
    1 & 0 & 0 & 1\\
    1 & 1 & 0 & 0\\
    1 & 1 & 1 & 1
    \end{pmatrix}\leq \begin{pmatrix}
    0 & 0 & 1 & 2\\
    1 & 0 & 1 & 1\\
    1 & 1 & 0 & 0\\
    1 & 1 & 1 & 1
    \end{pmatrix}=C'Q.$$
Using computer, 
    $$\rho(C)\approx 2.234\not\leq 2.148\approx \rho(C').$$
    This is because $C'$ is not $k$-rooted as $c'_{33}+c'_{34}=0\not\geq 1=c'_{43}+c'_{44}$.

@@

\bibitem{prn_fros2}
R. A. Horn , C. R. Johnson, {\it Matrix analysis}, Cambrigde University Press, 1985.
\end{thebibliography}

>>>>

In \cite{chang}, Cheng and Weng give many bounds of the spectral radius of a nonnegative square matrix. And based on their theory and Perron-Frobenius theorem, we give another approach to obtain an upper bound of the spectral radius and apply it on the adjacency matrix of a directed graph.
All theorems come from continuous discussions between C.W. Weng and K.H. Chen. These were all documented.\cite{src_files}

@@

\begin{proof}
    Multiplying the nonnegative vector $u$ in assumption~\ref{pre_thm_em1}

@@

%edges

\draw[edge] (1) to[bend right] (3);
\draw[edge] (1) to[bend right] (4);
\draw[edge] (2) to (1);
\draw[edge] (2) to[bend right] (4);
\draw[edge] (3) to (2);
\draw[edge] (3) to[bend right] (1);
\draw[edge] (3) to[bend right] (4); 
\draw[edge] (4)[bend right] to (1); 
\draw[edge] (4)[bend right] to (2); 
\draw[edge] (4)[bend right] to (3); 
\end{tikzpicture}
\end{center}

@@

%edges

\draw[edge] (1) to[bend right] (3);
\draw[edge] (1) to[bend right] (4);
\draw[edge] (2) to (1);
\draw[edge] (2) to (3);
\draw[edge] (3) to[bend right] (1);
\draw[edge] (3) to (2);
\draw[edge] (3) to[bend right] (4);
\draw[edge] (4)[bend right] to (1);
\draw[edge] (4) to (2);
\draw[edge] (4)[bend right] to (3);
\end{tikzpicture}
\end{center}

Note that both $G$ and $G'$ have the same number of edges, 
neither $A$ majories $A'$ nor $A'$ majories $A$. Our main result still can do comparison between 
$\rho(A)$ and $\rho(A')$. 
We first observe that $A$ is irreducible and $A'$ is $k$-rooted for $k=3$. Moreover 
$A'Q=\begin{pmatrix}
    0 & 0 & 1 & 2\\
    1 & 0 & 1 & 1\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 1
    \end{pmatrix}\geq \begin{pmatrix}
    0 & 0 & 1 & 2\\
    1 & 0 & 0 & 1\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 1
    \end{pmatrix}=AQ$. 
Hence Hence $\rho(A)\leq\rho(A')$ by Theorem~\ref{thm:conclusion} with $C=A$ and $C'=A'$. 
Both of the values of $\rho(C)$ and $\rho(C')$ are close to $2.511547$ by  calculating on computer. \cite[sage]{sage}
\end{exam}

@@

\begin{exam}
    Consider the following two $4\times 4$ matrices
    $$C=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 0 & 1\\
    1 & 1 & 0 & 0\\
    1 & 1 & 1 & 0
    \end{pmatrix},\quad C'=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 1 &  0\\
    1 & 1 & 0 & 0\\
    1 & 1 & 1 & 0
    \end{pmatrix}.$$
    With $n=4$ and $k=3$, we have  
    $$CQ=\begin{pmatrix}
    0 & 0 & 1 & 2\\
    1 & 0 & 0 & 1\\
    1 & 1 & 0 & 0\\
    1 & 1 & 1 & 1
    \end{pmatrix}\leq \begin{pmatrix}
    0 & 0 & 1 & 2\\
    1 & 0 & 1 & 1\\
    1 & 1 & 0 & 0\\
    1 & 1 & 1 & 1
    \end{pmatrix}=C'Q.$$
Using computer, \cite[sage]{sage}
    $$\rho(C)\approx 2.234\not\leq 2.148\approx \rho(C').$$
    This is because $C'$ is not $k$-rooted as $c'_{33}+c'_{34}=0\not\geq 1=c'_{43}+c'_{44}$.

All theorems come from continuous discussions between C.W. Weng and K.H. Chen. These were all documented.\cite{src_files}

@@

\bibitem{spec_rad}
A. E. Brouwer, W. H. Haemers, {\it Spectra of graphs}, Springer, 2012
\bibitem{prn_fros2}
R. A. Horn , C. R. Johnson, {\it Matrix analysis}, Cambrigde University Press, 1985.
\bibitem{sage}
[Sage] SageMath, the Sage Mathematics Software System (Version 9.2),
       The Sage Developers, 2020, https://www.sagemath.org.  %sage tutorial
\bibitem{src_files}
[github] K. H. Chen, C. W. Weng, source files to the essay, A simple method on comparison between spectral radii of two directed graphs,
     2021, https://reurl.cc/6yy32Z. % https://github.com/Urbaner3/NCTU_weng.

\end{thebibliography}

### 2021-01-20 spectral radius of matrix new method/thesis/article121420.tex
0121 final check and sending

Given a directed graph $G$, the spectral radius of $G$ is the spectral radius of the adjacency matrix of $G$, denoted by $\rho(G)$. Note that the spectral radius $\rho(G)$ is independent of the ordering of the vertex set of $G$.

@@

\subsection{Example}
    For the following $4\times 4$ matrix
    $$C=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 0 & 1\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 0
    \end{pmatrix},$$
    and corresponding directed graph, \cite[sage]{sage}
    \begin{center}
    \includegraphics{graph_C.PNG}
    \end{center}
    we choose the matrix and fix $k = 3$
    $$C'=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 1 & 0\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 0
    \end{pmatrix},$$
    Based on condition of theorem \ref{thm:conclusion},
    \begin{enumerate}[label=(\roman*)]
        \item  $C[[n]|[n-1]]\leq C'[[n]|[n-1]]$ and $c_{ik}+c_{in}\leq c'_{ik}+c'_{in}$ for all $1\leq i\leq n$;
        \item  there exists a $K$-rooted vector $v'=(v'_1, v'_2, \ldots, v'_n)^T$ and a scalar $\lambda'\in \mathbb{R}$
        such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$;
        \item there exists a nonnegative vector $v^T=(v_1, v_2, \ldots, v_n)$ and a scalar $\lambda\in \mathbb{R}$ 
        such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$;
        \item $v^Tv'>0.$
    \end{enumerate}
    Here $n=4,k=3$ so rewrite the condition of theorem \ref{thm_main} as following:
    \begin{enumerate}[label=(\roman*)]
        \item  $C[[4]|[3]]\leq C'[[4]|[3]]$ and $c_{i3}+c_{i4}\leq c'_{i3}+c'_{i4}$ for all $1\leq i\leq 4$;
        \item  there exists a $K$-rooted vector $v'=(v'_1, v'_2, v'_3, v'_4)^T$ and a scalar $\lambda'\in \mathbb{R}$
        such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$;
        \item there exists a nonnegative vector $v^T=(v_1, v_2, v_3, v_4)$ and a scalar $\lambda\in \mathbb{R}$ 
        such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$;
        \item $v^Tv'>0.$
    \end{enumerate}
    Then we check these conditions one by one,
    \begin{enumerate}[label=(\roman*)]
    \item 
    $$C[[4]|[3]]=\begin{pmatrix}
        0 & 0 & 1\\
        1 & 0 & 0\\
        1 & 1 & 0\\
        1 & 1 & 1
        \end{pmatrix},$$
    $$C'[[4]|[3]]=\begin{pmatrix}
        0 & 0 & 1\\
        1 & 0 & 1\\
        1 & 1 & 0\\
        1 & 1 & 1
        \end{pmatrix},$$
    thus,$C[[4]|[3]]\leq C'[[4]|[3]].$
    Write $c_{i3}+c_{i4}$ and $c'_{i3}+c'_{i4}$, for $1\leq i\leq 4$, as two row vectors $v_c$ and $v_{c'}$.
    $v_c = (s_1, s_2, s_3, s_4), s_i = c_{i3}+c_{i4},$ so $v_c = (2,1,1,1)$
    $v'_c = (s'_1, s'_2, s'_3, s'_4), s'_i = c'_{i3}+c'_{i4},$ so $v'_c = (2,1,1,1),$
    so $c_{i3}+c_{i4} \leq c'_{i3}+c'_{i4}.$
    \item 
    Let $\lambda'\in \mathbb{R}$ such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$.
    Then $\lambda' = 2.511547141694532$ and associated eigenvector $v'=(1, 0.8981609516297208, 1.255773570847266, 1.255773570847266)^T$.
    \item 
    Let  $\lambda\in \mathbb{R}$ such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$.
    Then $\lambda = 2.511547141694532$ and associated left eigenvector $v^T=(1, 0.715225238435091, 1, 0.7963219032594416)$.
    \item 
    $v^T * v' = (1, 0.715225238435091, 1, 0.7963219032594416) \\
    * (1, 0.8981609516297208, 1.255773570847266, 1.255773570847266)^T
    \\ = 3.89816095162972 >0.$

@@

    These facts satisfy the assumptions in Theorem~\ref{thm:conclusion}. Hence $\rho(C)\leq\rho(C')$. Actually, both of
     the values of $\rho(C)$ and $\rho(C')$ are close to 2.511547, and we have $\rho(C')-\rho(C)>0$ by calculating on
     computer.\cite[sage]{sage}
    Moreover, the corresponding directed graph of matrix $C'$ is, \cite[sage]{sage}
    \begin{center}
    \includegraphics{graph_Cprime.PNG}
    \end{center}

@@

\bibitem{prn_fros2}
R. A. Horn , C. R. Johnson, {\it Matrix analysis}, Cambrigde University Press, 1985.

\bibitem{sage}
[Sage] SageMath, the Sage Mathematics Software System (Version 9.2),
       The Sage Developers, 2020, https://www.sagemath.org.  %sage tutorial

>>>>

Given a directed graph $G$, the spectral radius of $G$ ,\cite{spec_rad}, is the
 spectral radius of the adjacency matrix of $G$, denoted by
 $\rho(G)$. Note that the spectral radius $\rho(G)$ is
 independent of the ordering of the vertex set of $G$.

@@

\subsection{Example}
    For the following $4\times 4$ matrix
    $$C=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 0 & 1\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 0
    \end{pmatrix},$$
    and corresponding directed graph, \cite[sage]{sage}
    \begin{center}
    \includegraphics{graph_C.PNG}
    \end{center}
    we choose the matrix and fix $k = 3$
    $$C'=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 1 & 0\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 0
    \end{pmatrix},$$

    which satisfies the assumptions in Theorem~\ref{thm:conclusion}. Hence $\rho(C)\leq\rho(C')$.
     Actually, both of the values of $\rho(C)$ and $\rho(C')$ are close to 2.511547, and we have
      $\rho(C')-\rho(C)>0$ by calculating on computer.

    Based on condition of theorem \ref{thm:conclusion},
    \begin{enumerate}[label=(\roman*)]
        \item  $C[[n]|[n-1]]\leq C'[[n]|[n-1]]$ and $c_{ik}+c_{in}\leq c'_{ik}+c'_{in}$ for all $1\leq i\leq n$;
        \item  there exists a $K$-rooted vector $v'=(v'_1, v'_2, \ldots, v'_n)^T$ and a scalar $\lambda'\in \mathbb{R}$
        such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$;
        \item there exists a nonnegative vector $v^T=(v_1, v_2, \ldots, v_n)$ and a scalar $\lambda\in \mathbb{R}$ 
        such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$;
        \item $v^Tv'>0.$
    \end{enumerate}
    Here $n=4,k=3$ so rewrite the condition of theorem \ref{thm_main} as following:
    \begin{enumerate}[label=(\roman*)]
        \item  $C[[4]|[3]]\leq C'[[4]|[3]]$ and $c_{i3}+c_{i4}\leq c'_{i3}+c'_{i4}$ for all $1\leq i\leq 4$;
        \item  there exists a $K$-rooted vector $v'=(v'_1, v'_2, v'_3, v'_4)^T$ and a scalar $\lambda'\in \mathbb{R}$
        such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$;
        \item there exists a nonnegative vector $v^T=(v_1, v_2, v_3, v_4)$ and a scalar $\lambda\in \mathbb{R}$ 
        such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$;
        \item $v^Tv'>0.$
    \end{enumerate}
    Then we check these conditions one by one,
    \begin{enumerate}[label=(\roman*)]
    \item 
    $$C[[4]|[3]]=\begin{pmatrix}
        0 & 0 & 1\\
        1 & 0 & 0\\
        1 & 1 & 0\\
        1 & 1 & 1
        \end{pmatrix},$$
    $$C'[[4]|[3]]=\begin{pmatrix}
        0 & 0 & 1\\
        1 & 0 & 1\\
        1 & 1 & 0\\
        1 & 1 & 1
        \end{pmatrix},$$
    thus,$C[[4]|[3]]\leq C'[[4]|[3]].$
    Write $c_{i3}+c_{i4}$ and $c'_{i3}+c'_{i4}$, for $1\leq i\leq 4$, as two row vectors $v_c$ and $v_{c'}$.
    $v_c = (s_1, s_2, s_3, s_4), s_i = c_{i3}+c_{i4},$ so $v_c = (2,1,1,1)$
    $v'_c = (s'_1, s'_2, s'_3, s'_4), s'_i = c'_{i3}+c'_{i4},$ so $v'_c = (2,1,1,1),$
    so $c_{i3}+c_{i4} \leq c'_{i3}+c'_{i4}.$
    \item 
    Let $\lambda'\in \mathbb{R}$ such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$.
    Then $\lambda' = 2.511547141694532$ and associated eigenvector $v'=(1, 0.8981609516297208, 1.255773570847266, 1.255773570847266)^T$.
    \item 
    Let  $\lambda\in \mathbb{R}$ such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$.
    Then $\lambda = 2.511547141694532$ and associated left eigenvector $v^T=(1, 0.715225238435091, 1, 0.7963219032594416)$.
    \item 
    $v^T * v' = (1, 0.715225238435091, 1, 0.7963219032594416) \\
    * (1, 0.8981609516297208, 1.255773570847266, 1.255773570847266)^T
    \\ = 3.89816095162972 >0.$

@@

    These facts satisfy the assumptions in Theorem~\ref{thm:conclusion}. Hence $\rho(C)\leq\rho(C')$. Actually, both of
     the values of $\rho(C)$ and $\rho(C')$ are close to 2.511547, and we have $\rho(C')-\rho(C)>0$ by calculating on
     computer.\cite[sage]{sage}

    Moreover, the corresponding directed graph of matrix $C'$ is, \cite[sage]{sage}
    \begin{center}
    \includegraphics{graph_Cprime.PNG}
    \end{center}

@@

\bibitem{prn_fros2}
R. A. Horn , C. R. Johnson, {\it Matrix analysis}, Cambrigde University Press, 1985.
\bibitem{sage}
[Sage] SageMath, the Sage Mathematics Software System (Version 9.2),
       The Sage Developers, 2020, https://www.sagemath.org.  %sage tutorial
